---
title: "The Fall album lyrics wordcloud"
author: "TP"
date: "02.01.2015"
output: html_document
---

# The Fall album lyrics wordcloud

## Summary wordcloud of pooled albums or singles/EP lyrics.

First we load libraries.
```{r loadlibs, message=FALSE}
library("rvest") # Simple web scraping for R
library("magrittr") # Piping: %>% operator (Ceci n'est pas un pipe)
library("tm") # A framework for text mining applications within R
library("wordcloud") # Word Clouds. Pretty ...
```

Then, we download links to albums and singles/EP from the great [The Fall online](http://thefall.org/discography/albums.html) discography section.

```{r discolinks}
discography <- list(albums="http://thefall.org/discography/albums.html",singles="http://thefall.org/discography/singles.html")
```

Function to download lyrics and munge html source using functions from 'rvest' package.

```{r downloadlyrics}
getlyrics <- . %>%
  html %>%
  html_nodes("a") %>% 
  html_attr("href") %>% {
    d <- grep("data",.)
    .[d]
  } %>% 
  file.path("http://thefall.org/discography",.) %>%
  sapply(html) %>% 
  lapply({. %>% html_nodes("i") %>% html_text(trim = TRUE)}) 
lyrics <- discography %>% lapply(getlyrics)
```

Function to download and munge album names.

```{r downloadtitles}
getTitles <- . %>%
  html %>%
  html_nodes("b:nth-child(3)") %>% 
  html_text(trim = TRUE)
Titles <- discography %>% lapply(getTitles)
```

Downloaded lyrics need some cleaning: replace weird apostrophes, remove line breaks. Then documents are turned to corpus using 'tm' package functions. 

```{r mungelyricstext}
replaceApost <- . %>% 
lapply({. %>% sapply({. %>% iconv(.,"","ASCII", "byte") %>% # convert to ascii to make apostrophes accessible via regex
                        gsub("<c3><a2><c2><80><c2><99>","'",.) %>% 
                        gsub("<c3><a2><c2><80><c2><98>","",.)})})
removeBreaks <- . %>% lapply(function(x) gsub("\n"," ",x)) %>% 
  lapply({. %>% unlist %>% paste(.,collapse = " ")})
mungeCorpus <- . %>% VectorSource %>% 
  Corpus %>% 
  TermDocumentMatrix(control = list(removePunctuation = TRUE, stopwords = TRUE, removeNumbers = TRUE)) %>% 
  as.matrix
# use these functions to clean downloaded lyrics and create text corpus
lyrics %<>% lapply(removeBreaks) %>% 
  lapply(replaceApost) %>% 
  lapply(mungeCorpus)
# add Album names
colnames(lyrics[[1]]) <- Titles[[1]]
colnames(lyrics[[2]]) <- Titles[[2]]
```

Function to summarise data for summary (all albums together) wordcloud.

```{r sumcloud}
munge <- . %>% rowSums %>%
            sort(., decreasing=TRUE) %>% 
            data.frame %>%
{word <- rownames(.)
 cbind(word,.)} %>%
  set_names(c("word","freq"))
# wordcloud plot function adapted to my dataframe
MyWC <- function(x) wordcloud(x$word, x$freq, random.order=FALSE, colors=brewer.pal(8, "Dark2"))
```

Plot wordcloud of all 30 'The Fall' albums.
```{r albumcloud, warning=FALSE}
lyrics[[1]] %>% munge %>% MyWC # Albums
```

Plot wordcloud of all 59 'The Fall' singles/EPs (not all of them have lyrics available).
```{r singlecloud, warning=FALSE}
lyrics[[2]] %>% munge %>% MyWC # Singles/EP
```

## Create wordclouds for each of 30 albums separately

Munge each album.

```{r albumsmunge}
Albums <- lyrics[[1]] %>% apply(.,2,{. %>% sort(decreasing=TRUE) %>% {
  word <- names(.)
  data.frame(., word)} %>% 
    set_names(c("freq","word"))})

```

Create short titles for album clouds. 

```{r albumtitles}
inf <- names(Albums) %>% {m <- regexpr("[0-9]{4}",.)
                          Year <- regmatches(., m)
                          Title <- gsub("[ ]?[[:punct:]0-9]{6}","",.)
                          data.frame(Year,Title)}
# function to shorten album titles
trimTitles <- function(titles) { # function nicked from stackoverflow
  len <- nchar(titles)
  maxlength <- 18
  cuts <- sapply(gregexpr(" ", titles), function(X) {
    max(X[X<maxlength])})
  titles[len>=maxlength] <- paste0(substr(titles[len>=maxlength], 0, cuts[len>=maxlength]), "..")
  titles
}
# add short titles to dataframe
inf$Short <- inf$Title %>% as.character %>% trimTitles
inf$num <- seq(1:30) # necessary to keep the albums order after split
inf <- split(inf, inf$num)
inf %<>% lapply(droplevels) %>% lapply({. %>% .[,-4]})
```

Merge album lyrics and title metadata.

```{r mergestuff}
Albums <- mapply(c, Albums, inf, SIMPLIFY = FALSE)
```


Modified wordcloud plot function with short titles perk.

```{r MyWCTits}
MyWCTits <- function(x) {
  wordcloud(x$word, x$freq, random.order=FALSE, colors=brewer.pal(8, "Dark2"))
  text(x=0.5, y=1.05, x$Short, vfont=c("serif","bold"), cex=1)}
```

Plot matrix of wordcloud plots.

```{r wordclouds, warning=FALSE, message=FALSE, fig.width=8, fig.height=8, results='hide'}
par(mfrow=c(5,6))
Albums %>% lapply(MyWCTits)
```
